\begin{thebibliography}{99}

\bibitem{CanadaWeb}
Government of Canada.
Canadian Centre for Cyber Security.
Retrieved July 26, 2021, from \url{https://cyber.gc.ca/en/guidance/cyber-threat-and-cyber-threat-actors}

\bibitem{MitreWeb}
MITRE.
MITRE ATT\&CK®.
Retrieved July 26, 2021, from \url{https://attack.mitre.org/}

\bibitem{Mitre}
MITRE ATT\&CK®.
Persistence, Tactic TA0003 - Enterprise.
Retrieved July 26, 2021, from \url{https://attack.mitre.org/tactics/TA0003/}

\bibitem{RTBook}
Eduardo Arriols. 
\emph{Chief Information Security Office: El Red Team de la empresa}. 
Vol1, 0xw0rd, 2018

\bibitem{PayloadAllTheThings}
Github - swisskyrepo/PayloadsAllTheThings.
Payload All the Things - Windows - Persistence.
Retrieved July 27, 2021, from \url{https://github.com/swisskyrepo/PayloadsAllTheThings/blob/master/Methodology\%20and\%20Resources/Windows\%20-\%20Persistence.md}

\bibitem{AwesomeRT}
Github - yeyintminthuhtut/Awesome-Red-Teaming.
Awesome Red Teaming - List of Awesome Red Team / Red Teaming Resources.
Retrieved July 28, 2021, from \url{https://github.com/yeyintminthuhtut/Awesome-Red-Teaming}

\bibitem{DomPersistenceCommands}
Cas van Cooten.
Windows \& Active Directory Exploitation Cheat Sheet and Command Reference.
Retrieved November 14, 2021, from \url{https://casvancooten.com/posts/2020/11/windows-active-directory-exploitation-cheat-sheet-and-command-reference/#domain-persistence}

%\bibitem{CommandsList}
%Cas van Cooten.
%Windows \& Active Directory Exploitation Cheat Sheet and Command Reference
%Retrieved July 28, 2021, from \url{https://casvancooten.com/posts/2020/11/windows-active-directory-exploitation-cheat-sheet-and-command-reference/}

\bibitem{SharPersist}
FireEye.
SharPersist - Windows persistence toolkit written in C\#.
Retrieved July 27, 2021, from \url{https://www.fireeye.com/blog/threat-research/2019/09/sharpersist-windows-persistence-toolkit.html}

\bibitem{Metasploit}
Rapid7.
Metasploit | Penetration Testing Software, Pen Testing Security.
Retrieved July 27, 2021, from \url{https://www.metasploit.com/}

%% this project %%

\bibitem{ThisProjectGit}
Github - alicenara/PersistenceAutomation. 
Persistence Deployment Automation.
Retrieved January 17, 2022, from \url{https://github.com/alicenara/PersistenceAutomation}

\bibitem{PagePersonnel}
Page Personnel. 
Study of trends, profiles and salaries - Remuneration study 2021. 
Retrieved September 7, 2021, from \url{https://www.pagepersonnel.es/sites/pagepersonnel.es/files/estudio_remuneracion_2021_sp.pdf}

\bibitem{TarifaLuzHora}
TarifaLuzHora. 
Hourly price of electricity rate. 
Retrieved September 7, 2021, from \url{https://tarifaluzhora.es/?tarifa=pcb&fecha=07\%2F09\%2F2021}

\bibitem{InternetPrice}
Selectra. 
Different Internet fares. 
Retrieved September 7, 2021, from \url{https://selectra.es/internet-telefono/internet}

\bibitem{susMatrix}
Jordi Garcia, Helena García, David López, Fermín Sánchez, Eva Vidal, Marc Alier y Jose Cabré. 
\emph{La sostenibilidad en los proyectos de ingeniería}. 
2013

\bibitem{KeePassWeb}
KeePass.
KeePass Password Safe.
Retrieved November 14, 2021, from \url{https://keepass.info/}

\bibitem{TortoiseSVNWeb}
TortoiseSVN.
TortoiseSVN - the coolest interface to (Sub)version control.
Retrieved November 14, 2021, from \url{https://tortoisesvn.net/}

\bibitem{Tunna}
Github - SECFORCE/Tunna.
Tunna is a set of tools which will wrap and tunnel any TCP communication over HTTP.
Retrieved November 14, 2021, from \url{https://github.com/SECFORCE/Tunna}

\bibitem{ExpiredDomainsWeb}
Expired Domains.
Backorder Pending Delete Domains.
Retrieved November 14, 2021, from \url{https://www.expireddomains.net/expired-domains/}

\bibitem{PaloAltoFilteringWeb}
Palo Alto Networks.
Palo Alto Networks URL filtering - Test A Site.
Retrieved November 14, 2021, from \url{https://urlfiltering.paloaltonetworks.com/}

\bibitem{DomainHunterWeb}
Github - threatexpress/domainhunter.
Domain Hunter - Checks expired domains for categorization/reputation.
Retrieved November 14, 2021, from \url{https://github.com/threatexpress/domainhunter}

\bibitem{Dridex}
Cyberbit.
How does Dridex gain persistency.
Retrieved November 8, 2021, from \url{https://www.cyberbit.com/blog/endpoint-security/how-does-dridex-gain-persistency/}

\bibitem{WindowsServices}
Microsoft - Microsoft Docs.
Services.
Retrieved November 8, 2021, from \url{https://docs.microsoft.com/en-us/previous-versions/windows/it-pro/windows-server-2008-R2-and-2008/cc772408(v=ws.11)}

\bibitem{WMIPersistence}
Penetration Testing Lab - Blog.
Persistence – WMI Event Subscription.
Retrieved November 8, 2021, from \url{https://pentestlab.blog/2020/01/21/persistence-wmi-event-subscription/}

\bibitem{COMObjects}
Penetration Testing Lab - Blog.
Persistence – COM Hijacking.
Retrieved November 9, 2021, from \url{https://pentestlab.blog/2020/05/20/persistence-com-hijacking/}

\bibitem{Shortcut}
Penetration Testing Lab - Blog.
Persistence – Shortcut Modification.
Retrieved October 18, 2021, from \url{https://pentestlab.blog/2019/10/08/persistence-shortcut-modification}

\bibitem{ForbixLNK}
Persianov on Security.
Windows worms. Forbix worm analysis.
Retrieved November 9, 2021, from \url{https://persianov.net/windows-worms-forbix-worm-analysis}

\bibitem{FireEyeWeb}
FireEye.
Cyber Security Experts \& Solution Providers.
Retrieved November 14, 2021, from \url{https://www.fireeye.com/}

\bibitem{Rapid7Web}
Rapid7.
Cybersecurity \& Compliance Solutions \& Services.
Retrieved November 14, 2021, from \url{https://www.rapid7.com/}

\bibitem{Meterpreter}
Offensive Security.
Understanding the Metasploit Meterpreter - Persistence.
Retrieved July 26, 2021, from \url{https://www.offensive-security.com/metasploit-unleashed/meterpreter-service/}

\bibitem{CobaltStrike}
Cobalt Strike.
Adversary Simulation and Red Team Operations Software.
Retrieved November 21, 2021, from \url{https://www.cobaltstrike.com/}

\bibitem{Mimikatz}
Github - gentilkiwi/mimikatz.
mimikatz - A little tool to play with Windows security.
Retrieved November 13, 2021, from \url{https://github.com/gentilkiwi/mimikatz}

\bibitem{Nishang}
Github - samratashok/nishang. 
Nishang - Offensive PowerShell for red team, penetration testing and offensive security.
Retrieved November 9, 2021, from \url{https://github.com/samratashok/nishang/blob/master/Utility/Add-Persistence.ps1}

\bibitem{Empire}
PowerShell Empire. 
Persistence.
Retrieved November 9, 2021, from \url{http://www.powershellempire.com/?page_id=139}

\bibitem{OSMarketShare}
Statcounter Global Stats. 
Desktop Operating System Market Share Worldwide.
Retrieved December 24, 2021, from \url{https://gs.statcounter.com/os-market-share/desktop/worldwide}


\bibitem{LinuxStatistics}
Security Space. 
OS/Linux Distributions using Apache.
Retrieved November 9, 2021, from \url{https://secure1.securityspace.com/s_survey/data/man.202110/apacheos.html}

\bibitem{MitreRC}
MITRE ATT\&CK®.
Boot or Logon Initialization Scripts: RC Scripts, Sub-technique T1037.004 -Enterprise.
Retrieved November 10, 2021, from \url{https://attack.mitre.org/techniques/T1037/004/}

\bibitem{MitreShells}
MITRE ATT\&CK®.
Event Triggered Execution: Unix Shell Configuration Modification, Sub-technique T1546.004 - Enterprise.
Retrieved November 10, 2021, from \url{https://attack.mitre.org/techniques/T1546/004/}

\bibitem{MitreSystemD}
MITRE ATT\&CK®.
Create or Modify System Process: Systemd Service, Sub-technique T1543.002 - Enterprise.
Retrieved November 11, 2021, from \url{https://attack.mitre.org/techniques/T1543/002/}

\bibitem{SystemDCode}
Hackers Vanguard.
Establishing Persistence with systemd.timers.
Retrieved November 11, 2021, from \url{https://hackersvanguard.com/establishing-persistence-systemd-timers/}

\bibitem{RedGhost}
Github - d4rk007/RedGhost. 
RedGhost - Linux post exploitation framework.
Retrieved November 11, 2021, from \url{https://github.com/d4rk007/redghost}

\bibitem{Linper}
Github - montysecurity/linper. 
linper - Linux Persistence Toolkit.
Retrieved November 11, 2021, from \url{https://github.com/montysecurity/linper}

\bibitem{MitrePreOS}
MITRE ATT\&CK®.
Pre-OS Boot, Technique T1542 - Enterprise.	
Retrieved December 24, 2021, from \url{https://attack.mitre.org/techniques/T1542/}

\bibitem{PuttyWeb}
PuTTY.
Download PuTTY - a free SSH and telnet client for Windows.
Retrieved November 14, 2021, from \url{https://www.putty.org/}

\bibitem{ProxyChainsWeb}
Github - haad/proxychains.
ProxyChains -  a tool that forces any TCP connection made by any given application to follow through proxy.
Retrieved November 14, 2021, from \url{https://github.com/haad/proxychains}

\bibitem{PoshC2}
Github - nettitude/PoshC2.
PoshC2 - A proxy aware C2 framework used to aid red teamers with post-exploitation and lateral movement.	
Retrieved December 24, 2021, from \url{https://github.com/nettitude/PoshC2}

\bibitem{newtrixBook}
Brian Svidergol. 
\emph{What is Active Directory?}. 
netwrix, 2019

\bibitem{MitreNTDS}
MITRE ATT\&CK®.
OS Credential Dumping: NTDS, Sub-technique T1003.003 - Enterprise.	
Retrieved December 25, 2021, from \url{https://attack.mitre.org/techniques/T1003/003/}

\bibitem{GoldenTicket1}
A paper by Chun Feng, Tal Be’ery and Stewart McIntyre.
Domain Persistence: Golden Ticket Attack.
Retrieved November 14, 2021, from \url{https://www.hackingarticles.in/domain-persistence-golden-ticket-attack/}

\bibitem{Tarlogic}
Tarlogic.
Kerberos (I): ¿Cómo funciona Kerberos? - Teoría.
Retrieved November 21, 2021, from \url{https://www.tarlogic.com/es/blog/como-funciona-kerberos/}


\bibitem{BloodHound}
Github - BloodHoundAD/BloodHound. 
BloodHound - Six Degrees of Domain Admin.
Retrieved November 12, 2021, from \url{https://github.com/BloodHoundAD/BloodHound}

\bibitem{SharpHound}
BloodHound. 
SharpHound — BloodHound 3.0.3 documentation.
Retrieved November 13, 2021, from \url{https://bloodhound.readthedocs.io/en/latest/data-collection/sharphound.html}

\bibitem{ADExplorer}
Microsoft Docs. 
AD Explorer - Windows Sysinternals.
Retrieved November 13, 2021, from \url{https://docs.microsoft.com/en-us/sysinternals/downloads/adexplorer}


\bibitem{Rubeus}
Github - GhostPack/Rubeus.
Rubeus - Trying to tame the three-headed dog.
Retrieved November 14, 2021, from \url{https://github.com/GhostPack/Rubeus}

\bibitem{Impacket}
Github - SecureAuthCorp/impacket.
impacket - Impacket is a collection of Python classes for working with network protocols.
Retrieved November 14, 2021, from \url{https://github.com/SecureAuthCorp/impacket}

\bibitem{PowerView}
Github - PowerShellMafia/PowerSploit.
PowerSploit - PowerView.
Retrieved November 14, 2021, from \url{https://github.com/PowerShellMafia/PowerSploit/tree/master/Recon}

\bibitem{MitreDisc}
MITRE ATT\&CK®.
Discovery, Tactic TA0007 - Enterprise.
Retrieved November 14, 2021, from \url{https://attack.mitre.org/tactics/TA0007/}

\bibitem{MitreLM}
MITRE ATT\&CK®.
Lateral Movement, Tactic TA0008 - Enterprise.
Retrieved November 14, 2021, from \url{https://attack.mitre.org/tactics/TA0008/}

\bibitem{MitrePE}
MITRE ATT\&CK®.
Privilege Escalation, Tactic TA0004 - Enterprise.
Retrieved November 14, 2021, from \url{https://attack.mitre.org/tactics/TA0004/}


\bibitem{GoldenTicket2}
Hacking Articles - Raj Chandel's Blog.
Golden Ticket.
Retrieved November 14, 2021, from \url{https://pentestlab.blog/2018/04/09/golden-ticket/}

\bibitem{GPOAttack}
Active Directory Security.
Sneaky Active Directory Persistence \#17: Group Policy.
Retrieved November 14, 2021, from \url{https://adsecurity.org/?p=2716}

\bibitem{MitreGPOs}
MITRE ATT\&CK®.
Boot or Logon Initialization Scripts: Network Logon Script, Sub-technique T1037.003 - Enterprise.	
Retrieved December 25, 2021, from \url{https://attack.mitre.org/techniques/T1037/003/}

\bibitem{SkeletonKey}
A paper by Chun Feng, Tal Be’ery and Stewart McIntyre.
\emph{DIGITAL "BIAN LIAN" (FACE CHANGING): THE SKELETON KEY MALWARE}.
Retrieved November 14, 2021, from \url{https://www.virusbulletin.com/uploads/pdf/magazine/2016/vb201601-skeleton-key.pdf}

\bibitem{AttackingAD}
Gitlab - zer1t0.
Attacking Active Directory: 0 to 0.9.
Retrieved November 14, 2021, from \url{https://zer1t0.gitlab.io/posts/attacking_ad}

\bibitem{DomPersistenceExplained}
HackTricks.
Active Directory Methodology. 
Retrieved November 14, 2021, from \url{https://book.hacktricks.xyz/windows/active-directory-methodology#persistence}

\bibitem{WikiScript}
Wikipedia.
Scripting language.	
Retrieved January 17, 2022, from \url{https://en.wikipedia.org/wiki/Scripting_language}

%% dev %%

\bibitem{Mistica}
Github - IncideDigital/Mistica.
Mistica - An open source swiss army knife for arbitrary communication over application protocols.
Retrieved November 14, 2021, from \url{https://github.com/IncideDigital/Mistica}

\bibitem{HTTPRevshell}
Github - 3v4Si0N/HTTP-revshell.
Powershell HTTP/S Reverse Shell - Powershell reverse shell using HTTP/S protocol with AMSI bypass and Proxy Aware.
Retrieved November 14, 2021, from \url{https://github.com/3v4Si0N/HTTP-revshell}

%% conclusions %%

\bibitem{fail2ban}
Github - fail2ban/fail2ban.
fail2ban - Daemon to ban hosts that cause multiple authentication errors 
Retrieved November 20, 2021, from \url{https://github.com/fail2ban/fail2ban}

%% tools in the table %%

\bibitem{reGeorg}
Github - sensepost/reGeorg.
reGeorg - The successor to reDuh.
Retrieved November 14, 2021, from \url{https://github.com/sensepost/reGeorg}


\bibitem{pivotnacci}
Github - blackarrowsec/pivotnacci.
pivotnacci - A tool to make socks connections through HTTP agents.
Retrieved November 14, 2021, from \url{https://github.com/blackarrowsec/pivotnacci}

\bibitem{Merlin}
Github - Ne0nd0g/merlin.
Merlin is a cross-platform post-exploitation HTTP/2 Command \& Control server and agent written in golang.
Retrieved November 14, 2021, from \url{https://github.com/Ne0nd0g/merlin}


\bibitem{Dnscat2}
Github - iagox86/dnscat2.
Dnscat2, a DNS tunnel.
Retrieved November 14, 2021, from \url{https://github.com/iagox86/dnscat2}

\bibitem{icmpsh}
Github - hemp3l/icmpsh.
icmpsh - Simple Reverse ICMP Shell.
Retrieved November 14, 2021, from \url{https://github.com/hemp3l/icmpsh}

\bibitem{icmpshPython}
Github - bdamele/icmpsh.
icmpsh - a Python port of a simple reverse ICMP shell.
Retrieved November 14, 2021, from \url{https://github.com/bdamele/icmpsh}

\bibitem{Sliver}
Github - BishopFox/sliver.
Sliver - Adversary Emulation Framework.
Retrieved November 14, 2021, from \url{https://github.com/BishopFox/sliver}






%%%% Anterior





% a recolocar















%% others %%

% poner wiki como elemento?
%\bibitem{Wiki}
%Wikipedia. 
%Computer security - Wikipedia.
%Retrieved from \url{https://en.wikipedia.org/wiki/Computer_security}

%\bibitem{MalwarePersistenceCollection}
%Github - Karneades/malware-persistence. 
%Collection of malware persistence information.
%Retrieved September 14, 2021, from \url{https://github.com/Karneades/malware-persistence}

\end{thebibliography}